---
title: "Linear Regression"
author: "Hussain Hadah"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    css: ["default", "assets/sydney-fonts.css", "assets/sydney.css"]
    self_contained: false # if true, fonts will be stored locally
    seal: true # show a title slide with YAML information
    includes:
      in_header: "assets/mathjax-equation-numbers.html"
    nature:
      beforeInit: ["assets/remark-zoom.js", "https://platform.twitter.com/widgets.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9' # alternatives '16:9' or '4:3' or others e.g. 13:9
      navigation:
        scroll: false # disable slide transitions by scrolling
---
name: toc
layout: true
<div style="position: absolute;left:20px;bottom:5px;color:black;font-size: 12px;">`r rmarkdown::metadata$author` (UH) | `r rmarkdown::metadata$subtitle` | `r format(Sys.time(), '%d %B %Y')`</div>

<!--- `r rmarkdown::metadata$subtitle` | `r format(Sys.time(), '%d %B %Y')`-->

---

```{r, load_refs, echo=FALSE, cache=FALSE, message=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE, 
           bib.style = "authoryear", 
           cite.style = 'authoryear', 
           style = "markdown",
           hyperlink = FALSE, 
           dashed = FALSE)
myBib <- ReadBib("assets/example.bib", check = FALSE)
top_icon = function(x) {
  icons::icon_style(
    icons::fontawesome(x),
    position = "fixed", top = 10, right = 10
  )
}
```

```{r setup, include=FALSE}
# xaringanExtra::use_scribble() ## Draw on slides. Requires dev version of xaringanExtra.

options(htmltools.dir.version = FALSE)
library(knitr)
opts_chunk$set(
  fig.align="center",  
  fig.height=4, fig.width=6,
  out.width="748px", out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=F#, echo=F, warning=F, message=F
  )
```

```{r data, include=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, hrbrthemes, sysfonts, 
               showtext,  gghighlight)

load(file.path("/Users/hhadah/Documents/GiT/teaching-demo/linear-regression","parenthood.Rdata"))

font_add_google("Fira Sans", "firasans")
font_add_google("Fira Code", "firasans")

showtext_auto()
options(modelsummary_format_numeric_latex = "mathmode")
showtext_opts(dpi = 300)

theme_customs <- function() {
  theme_minimal(base_family = "IBM Plex Sans Condensed") +
    theme(panel.grid.minor = element_blank(),
          plot.background = element_rect(fill = "white", color = NA),
          plot.title = element_text(face = "bold", size = rel(2)),
          axis.title = element_text(face = "bold"),
          strip.text = element_text(face = "bold"),
          strip.background = element_rect(fill = "grey80", color = NA),
          legend.title = element_text(face = "bold", size = rel(1)),
          axis.text.y  = element_text(size = 18),
          axis.text.x  = element_text(size = 18),
          axis.title.x = element_text(size = 20),
          axis.title.y = element_text(size = 20),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.text = element_text(size = rel(1)))
}
```

```{css, echo=F}
    /* Table width = 100% max-width */

    .remark-slide table{
        width: 100%;
    }

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(2n) {
        background-color: white;
    }
    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }
```

## Table of contents

1. [Prologue](#prologue)

---
class: segue-red

# Prologue

---

## What We Will Learn

.blockquote[
### `r icons::fontawesome("location-arrow")` Aim

#### 1. Introduce Linear Regression 

#### 2. Understand the Intuition of Linear Regression 

#### 3. Deriving the Regression Equation

#### 4. Running a Linear Regression in R
]


---
class: segue-red

# Linear Regression

---
## What is a Linear Regression?

.Large[.content-box-purple[
A linear regression is a tool used by economists to analyze relationships between interval scale predictors and interval scale outcomes
]]

--
- What does this definition reminde you of?

--
- Correlations

---
## What is a linear regression model?

```{r regression0, fig.cap="Scatterplot showing grumpiness as a function of hours slept.", echo=FALSE, warning=FALSE}

parenthood |> 
  ggplot(aes(x = dan.sleep, y = dan.grump)) +
  geom_point() +
  theme_customs() +
  labs(x = "My sleep (hours)",
       y = "My grumpiness (0-100)")
```

---
count: false
## What is a linear regression model?

```{r regression2, fig.cap="Scatterplot showing grumpiness as a function of hours slept.", echo=FALSE, warning=FALSE, fig.height=3}

parenthood |> 
  ggplot(aes(x = dan.sleep, y = dan.grump)) +
  geom_point() +
  theme_customs() +
  theme(axis.title.y = element_text(size = 15)) +
  labs(x = "My sleep (hours)",
       y = "My grumpiness (0-100)")
```

- We can predict my grumpiness based on the data points

---
count: false

## What is a linear regression model?

```{r regression3, fig.cap="Scatterplot showing grumpiness as a function of hours slept.", echo=FALSE, warning=FALSE, fig.height=3}

parenthood |> 
  ggplot(aes(x = dan.sleep, y = dan.grump)) +
  geom_point() +
  geom_vline(xintercept = 7, color = "red", linetype = "dashed") +
  gghighlight(dan.sleep < 7.1 & dan.sleep >  6.9) + 
  theme_customs() +
  theme(axis.title.y = element_text(size = 15)) +
  labs(x = "My sleep (hours)",
       y = "My grumpiness (0-100)")
```

- We can predict my grumpiness based on the data points
- How grumpy could I be if I had 7 hours of sleep?

---
## Predict my grumpiness given any hours of sleep

```{r regression4, fig.cap="Scatterplot showing grumpiness as a function of hours slept.", echo=FALSE, warning=FALSE}

parenthood |> 
  ggplot(aes(x = dan.sleep, y = dan.grump)) +
  geom_point() +
  theme_customs() +
  labs(x = "My sleep (hours)",
       y = "My grumpiness (0-100)")
```


---
## Regression: “Best Fitting” Line Through Cloud of Points

```{r regression5, fig.cap="Scatterplot showing grumpiness as a function of hours slept.", echo=FALSE, warning=FALSE, results='hide', message = FALSE}

parenthood |> 
  ggplot(aes(x = dan.sleep, y = dan.grump)) +
  geom_point() +
  suppressMessages(geom_smooth(method = "lm", se = FALSE)) +
  theme_customs() +
  labs(x = "My sleep (hours)",
       y = "My grumpiness (0-100)")
```

---
class: segue-red

# Deriving the Regression Equation

---
## The Regression Equation

#### To derive the regression, let remember some highschool math

The formula of a straight line is given by:
$$
y = mx + c
$$
--

- There are two variables, $x$ and $y$

- There are two coefficients, $m$ and $c$

--
- $m$ is the slope of a line

- $c$ y-intercept 

```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}

# https://hhadah.github.io/MicroSlides/MyPresentations/intro/intro.html

infile = list.files(path = "/Users/hhadah/Documents/GiT/teaching-demo/index.html", pattern = 'index.html')
pagedown::chrome_print(input = infile, timeout = 100)
xaringan::inf_mr() 
```

